Motional Ground-State Cooling Outside the Lamb-Dicke Regime

### 2017-07-12 paper.tex
Clarify Lamb-Dicke regime

Prepare quantum ground-state for a variety of systems including
single atoms in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
has been successful using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems has all been applied in the Lamb-Dicke regime \fxnote{define}.
A challenging goal is to apply ground-state cooling to system such as atoms with light masses
or laser-cooled polar molecules, that are often outside of the Lamb-Dicke regime.
In this letter, we overcome this challenge and demonstrate cooling of single sodium atoms
that are trapped in optical tweezers to the motional ground state.
We achieve a single atom ground-state probability of $P_0=77(4)\%$ by utilizing cooling of
very high-order sidebands in a carefully optimized cooling sequence.
Our approach is  general and opens up ground-state cooling for other systems.

@@

Fortunately, the large Lamb-Dicke parameters also provide a way to overcome OP heating.
The Raman transitions in our configuration have
$\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{0.40(1), 0.341(2), 0.291(1)\}$,
which allow strong coupling to higher order sidebands for atoms in large motional states
as calculated in Figure~\ref{f-ld}\fxnote{A or B?}.x
To offset heating from OP initially, higher-order Raman cooling sidebands can be utilized
to remove more motional quanta in a single cooling pulse.
Since the coupling strengths of different orders do not reach minima for the same pulse duration,
using multiple orders of cooling sidebands would avoid accumulations of population
near the coupling minima.

@@

After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights to be $\bar n / (\bar n + 1)$,
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
Outside of the Lamb-Dickes regime, however,
the coupling strength deviates from this simple scaling rule rapidly
as shown in figure \ref{f-ld}A and since the Raman spectroscopy is performed
with a relatively large Raman Lamb-Dickes parameters $\eta^R$, we cannot use this formula to
measure $\bar n$. Instead, given the absence of the second order cooling sideband,
we assume that the atoms are colder than

@@

we therefore use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify the ground-state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)\%$ and $92(2)\%$ and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping and obtained \fxnote{70(?)} $\mu$K.

@@

Combining the axial and radial cooling results,
we have prepared a single Na atom with 3D ground state probability of $77(4)\%$.
We measuring a heating rate of the single atoms in the tweezer to be \fxnote{??}\%
due to trapping light scattering.\fxnote{ciet Grimm}
The ground-state preparation fidelity is currently limited by off-resonant scattering
from the Raman beams,
which are measured to be between $3$ to $15$ kHz for the detuning and power we use,
and the resonance shift caused by magnetic field fluctuations.
In addition, there are $15$\% population loss during the experiment,
including $5\%$ from single atom imaging and  $10\%$ very high-lying initial motional states
that are not efficiently cooled by RSC, but off-resonantly heat by the Raman beams.
We are planing to improve these by increasing the detuning of the Raman beams,
addressing wider range of anharmonicity by frequency sweep...,
implementing better control of the magnetic field and optimize the temperature during imaging.
We expect these changes to improve the cooling performance further.
Another improvement could be to implement grey molasses cooling to achieve
a lower starting temperature before RSC. \fxnote{cite}

>>>>

Prepare quantum ground-state for a variety of systems including
single atoms in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
has been successful using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems has all been applied in the Lamb-Dicke regime where
the size of the wavefunction is much smaller than the wavelengths of
the Raman and optical pumping light.
A challenging goal is to apply ground-state cooling to system such as atoms with light masses
or laser-cooled polar molecules, that are often outside of the Lamb-Dicke regime.
In this letter, we overcome this challenge and demonstrate cooling of single sodium atoms
that are trapped in optical tweezers to the motional ground state.
We achieve a single atom ground-state probability of $P_0=77(4)\%$ by utilizing cooling of
very high-order sidebands in a carefully optimized cooling sequence.
Our approach is  general and opens up ground-state cooling for other systems.

@@

Fortunately, the large Lamb-Dicke parameters also provide a way to overcome OP heating.
The Raman transitions in our configuration have
$\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{0.40(1), 0.341(2), 0.291(1)\}$,
which allow strong coupling to higher order sidebands for atoms in large motional states
as calculated in Figure~\ref{f-ld}\fxnote{A or B?}.
To offset heating from OP initially, higher-order Raman cooling sidebands can be utilized
to remove more motional quanta in a single cooling pulse.
Since the coupling strengths of different orders do not reach minima for the same pulse duration,
using multiple orders of cooling sidebands would avoid accumulations of population
near the coupling minima.

@@

After cooling, the first and second order cooling sidebands on both radial axes are suppressed.
It is important to note that the simple sideband thermometry formula,
which predicts the ratio of the cooling and heating sideband heights to be $\bar n / (\bar n + 1)$,
assumes that the coupling strength for different motional states to be proportional to $\sqrt{n}$.
Outside of the Lamb-Dickes regime, however,
the coupling strength deviates from this simple scaling rule rapidly
as shown in figure \ref{f-ld}A and since the Raman spectroscopy is performed
with a relatively large Raman Lamb-Dickes parameters $\eta^R$, we cannot use this formula to
measure $\bar n$. Instead, given the absence of the second order cooling sideband,
we assume that the atoms are cold enough to have stronger coupling to the first order cooling
sideband than the second order, in which case the

@@

we therefore use an independent measurement of Rabi flopping on the carrier (Fig. \ref{f-radial}B)
and first-order heating(Fig. \ref{f-radial}C) sidebands
to verify the ground-state population~\cite{Meekhof1996}.
The data is fitted using the same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)\%$ and $92(2)\%$ and show good agreement with
the population extracted from the sideband thermometry.
We also extracted the initial temperature before cooling
from the motional-state decohered Rabi flopping and obtained $70 \mu$K.

@@

Combining the axial and radial cooling results,
we have prepared a single Na atom with 3D ground state probability of $77(4)\%$.
We measuring a heating rate of the single atoms in the tweezer to be \fxnote{??}\%
due to trapping light scattering.\fxnote{cite Grimm}
The ground-state preparation fidelity is currently limited by off-resonant scattering
from the Raman beams,
which are measured to be between $3$ to $15$ kHz for the detuning and power we use,
and the resonance shift caused by magnetic field fluctuations.
In addition, there are $15$\% population loss during the experiment,
including $5\%$ from single atom imaging and  $10\%$ very high-lying initial motional states
that are not efficiently cooled by RSC, but off-resonantly heat by the Raman beams.
We are planing to improve these by increasing the detuning of the Raman beams,
addressing wider range of anharmonicity by frequency sweep...,
implementing better control of the magnetic field and optimize the temperature during imaging.
We expect these changes to improve the cooling performance further.
Another improvement could be to implement grey molasses cooling to achieve
a lower starting temperature before RSC. \fxnote{cite}
